Generalized Tractability

### 2023-04-02 GeneralizedTractability_230310.tex
Updates from Overleaf
Fred's Home Computer

For $\varepsilon$ taking on all positive values, $j(\varepsilon,\vp)$ takes on all values greater than or equal to $j(\infty,\vp)$, so
\[
j\le C_{\vp} T(\lambda_{j,d}^{-1},1,\vp)+1, \qquad \forall j \ge j(\infty,\vp).
\]
This implies via our technical assumption \eqref{eq:ptauassume} that
\begin{gather*}
 K_{\vp,\tau}\,T (\lambda_{j,d}^{-1},1,\tau \vp) \ge
 [T(\lambda_{j,d}^{-1},1, \vp)]^\tau
 \ge
  \left[\frac{(j-1)}{C_{\vp}}\right]^\tau \qquad \forall j \ge j(\infty,\vp) \\
 \frac{1}{T (\lambda_{j,d}^{-1},1,\tau \vp)} \le
\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(j-1)^\tau}\qquad \forall j \ge j(\infty,\vp)
\end{gather*}
Summing both sides of this last inequality for $j$ from $j(\infty,\vp) \ge 1$ to $\infty$ yields 
\begin{align*}
\sup_{d\in\naturals} \sum_{j=j (\infty,\vp)}^\infty \frac{1}{T (\lambda_{j,d}^{-1},1, \tau \vp)}
& \le  \sup_{d\in\naturals} \frac{1}{T (\lambda_{1,d}^{-1},1, \tau \vp)} + K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{j=2}^\infty \frac{1}{(j-1)^\tau} \\
& \le\frac{1}{T (\lambda_{\max}^{-1},1, \tau \vp)} +  K_{\vp,\tau}\, C_{\vp}^\tau
\zeta (\tau)  < \infty,
\end{align*}
where $\zeta$ denotes the Riemann zeta function.
This yields \eqref{eq:strong_tractiff3} with $\vp$ replaced by $\vp' = \tau \vp$ and $L_{\vp'} = j (\infty,\vp)$, and so we see the necessity of $\eqref{eq:strong_tractiff3}$.

@@

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & = T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}} \Biggr ] \\
       & \le T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]
        \\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp}\, T(0,d,\vp)}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right] \\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},d,\vp)} \,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},d,\vp)} \,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le \frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},d,\vp)}  T(\varepsilon^{-1},d,2\vp)  + S_{\vp}\, T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le \underbrace{\left[\frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]}_{=:C_{2\vp}} \,T(\varepsilon^{-1},d,2\vp)
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}. \\

@@

\fred{We need to check the proof for
\begin{itemize}
\item correctness
\item exponents
\item whether the additional assumptions are too restrictive and can be generalized
\item whether we need $\tau$ in the assumption
\item Do we need non-decreasing or strictly increasing?
\end{itemize}}
\begin{definition}[Weak Tractability]
If $n(\varepsilon,d)$ denotes the minimal number of function values needed to approximate all functions. We say that the problem is \emph{weakly tractable} if
    \[\lim_{\varepsilon^{-1}+d\rightarrow \infty} \frac{\ln n(\varepsilon,d)}{\varepsilon^{-1}+d} = 0\]
\end{definition}
\begin{definition}[Quasi-polynomial Tractability]
    The problem is \emph{quasi-polynomially tractable}, if there exist two constants $C, t> 0$ such that
    \[
    n(\varepsilon,d) \leq C\exp\{t(1+\ln(1/\varepsilon))(1+\ln(d))\}
    \] for all $\varepsilon \in (0,1)$ and $d\in\mathbb{N}$.
\end{definition}
\begin{thebibliography}{99}

\bibitem{NW08} E.~Novak,H.~Wo\'zniakowski. \textit{Tractability of Multivariate Problems, Volume I: Linear Information}.
European Mathematical Society Publishing House, Zurich, 2008.
 ×
\end{thebibliography}

>>>>

For $\varepsilon$ taking on all positive values, $j(\varepsilon,\vp)$ takes on all values greater than or equal to $j(\infty,\vp)$, so
\[
j\le C_{\vp} T(\lambda_{j,d}^{-1},1,\vp)+1, \qquad \forall j \ge j(\infty,\vp).
\]
This implies via our technical assumption \eqref{eq:ptauassume} that
\begin{gather*}
 K_{\vp,\tau}\,T (\lambda_{j,d}^{-1},1,\tau \vp) \ge
 [T(\lambda_{j,d}^{-1},1, \vp)]^\tau
 \ge
  \left[\frac{(j-1)}{C_{\vp}}\right]^\tau \qquad \forall j \ge j(\infty,\vp) \\
 \frac{1}{T (\lambda_{j,d}^{-1},1,\tau \vp)} \le
\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(j-1)^\tau}\qquad \forall j \ge j(\infty,\vp)
\end{gather*}
Summing both sides of this last inequality for $j$ from $j(\infty,\vp) \ge 1$ to $\vinfty$ yields 
\begin{align*}
\sup_{d\in\naturals} \sum_{j=j (\infty,\vp)}^\infty \frac{1}{T (\lambda_{j,d}^{-1},1, \tau \vp)}
& \le  \sup_{d\in\naturals} \frac{1}{T (\lambda_{1,d}^{-1},1, \tau \vp)} + K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{j=2}^\infty \frac{1}{(j-1)^\tau} \\
& \le\frac{1}{T (\lambda_{\max}^{-1},1, \tau \vp)} +  K_{\vp,\tau}\, C_{\vp}^\tau
\zeta (\tau)  < \infty,
\end{align*}
where $\zeta$ denotes the Riemann zeta function.
This yields \eqref{eq:strong_tractiff3} with $\vp$ replaced by $\vp' = \tau \vp$ and $L_{\vp'} = j (\infty,\vp)$, and so we see the necessity of $\eqref{eq:strong_tractiff3}$.

@@

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & = T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}} \Biggr ] \\
       & \le T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]
        \\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp}\, T(0,d,\vp)}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp}\right] \\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le \frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)}  T(\varepsilon^{-1},d,2\vp)  + S_{\vp}\, T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le \underbrace{\left[\frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp}\right]}_{=:C_{2\vp}} \,T(\varepsilon^{-1},d,2\vp)
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}. \\

@@

\fred{We need to check the proof for
\begin{itemize}
\item correctness
\item exponents
\item whether the additional assumptions are too restrictive and can be generalized
\item whether we need $\tau$ in the assumption
\item Do we need non-decreasing or strictly increasing?
\end{itemize}}
\begin{definition}[Weak Tractability]
If $n(\varepsilon,d)$ denotes the minimal number of function values needed to approximate all functions. We say that the problem is \emph{weakly tractable} if
    \[\lim_{\varepsilon^{-1}+d\rightarrow \infty} \frac{\ln n(\varepsilon,d)}{\varepsilon^{-1}+d} = 0\]
\end{definition}
\begin{definition}[Quasi-polynomial Tractability]
    The problem is \emph{quasi-polynomially tractable}, if there exist two constants $C, t> 0$ such that
    \[
    n(\varepsilon,d) \leq C\exp\{t(1+\ln(1/\varepsilon))(1+\ln(d))\}
    \] for all $\varepsilon \in (0,1)$ and $d\in\mathbb{N}$.
\end{definition}

\peter{
\section{Weak $T$-$(s,t)$-weak tractability (work in progress)} \label{sec:weak}

\begin{definition}[$T$-$(s,t)$-weak tractability]
Let $s,t>0$. We say that the problem is \emph{$T$-$(s,t)$ weakly tractable} if
    \[\lim_{\varepsilon^{-1}+d\rightarrow \infty} \frac{\log n(\varepsilon,d)}{T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp)} = 0.\]
\end{definition}

We show a necessary and sufficient condition for $T$-$(s,t)$-weak tractability, working analogously to \cite{WW17}.
\begin{theorem}\label{thm:weak}
For $c>0$, define
\[
\sigma (c,d,s,\vp):= \sum_{i=1}^\infty \exp \left(-c\, T(\lambda_{i,d}^{-s},1,\vp)\right).
\]
Then the problem is $T$-$(s,t)$-weakly tractable if and only if
\[
\mu (c,s,t,\vp)=\sup_{d\in\naturals} \sigma (c,d,s,\vp) \exp\left(-c\, T(0,d^t,p,q)\right)<\infty
\]
holds for all $c>0$.
\end{theorem}
\begin{proof}
 We first show the necessity of the condition in the theorem.   Indeed, suppose that the problem is  $T$-$(s,t)$-weakly tractable. Then, for any $c>0$ there exists a positive integer $V=V(c)$ such that  
 \[
 n(\varepsilon,d)\le 
 \left\lfloor \exp\left(c\,
 \left(T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp)\right)
 \right) \right\rfloor
 \]
 whenever $\varepsilon^{-1} + d \ge V$. For $d\in\naturals$ 
 choose $\varepsilon>0$ such that $\varepsilon^{-1}\ge \max\{1,V-d\}$. Since the eigenvalues $\lambda{i,d}$ are non-increasing in $i$, we know that
 \[
 \lambda_{\lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1}^2 \le \varepsilon^2
 \]
 for any $\varepsilon$ satisfying $\varepsilon^{-1}\ge \max\{1,V-d\}$.
Let 
\[
n=n(\varepsilon):= \lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1
\]
and
\[
n^*:= \lfloor \exp(c\,
 (T(\max\{1,(V-d)^s\},1,\vp) + T(0,d^t,\vp))) \rfloor+1.
\]
By varying $\varepsilon^{-1}$ ini the interval 
$[\max\{1,V-d\},\infty)$ we can make $n(\varepsilon)$ attain 
any integer value greater than or equal to $n^*$.

[TO BE CONTINUED]

\end{proof}
}

\begin{thebibliography}{99}

\bibitem{NW08} E.~Novak,H.~Wo\'zniakowski. \textit{Tractability of Multivariate Problems, Volume I: Linear Information}.
European Mathematical Society Publishing House, Zurich, 2008.
 \bibitem{WW17} A.G.~Werschulz,H.~Wo\'zniakowski. A new characterization of $(s,t)$-weak tractability. J. Complexity 38, 68--79, 2017. 
\end{thebibliography}
